---
title: "R for Manifold"
author: "Michael Sumner"
date: 2015-06-09
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```
# R for Manifold

Manifold GIS and R make for a powerful partnership, but the coupling between them has been relatively loose and sketchy. 

Two key recent R packages make the coupling more compelling:

* [wkb](http://cran.rstudio.com/web/packages/wkb/index.html): Convert Between Spatial Objects and Well-Known Binary (WKB) Geometry
* [rClr](https://rclr.codeplex.com)

The main ways we connect R and Manifold are 1) read drawing layers with data from Manifold .map files via SQL queries
2) drive the Manifold API directly via .Net. 

The first provides a pretty tight mapping of high-level data types, i.e. in Manifold we have a drawing and in R we have a Spatial layer and moving from one to the other is easy. 

The second provides a lot more power but we need to do more work to transfer data between the systems. 

There are lots of other pathways, including GDAL as a third/fourth party and via file transfer. 


## Examples

* illustrate this example in both  R and Manifold 

https://github.com/mdsumner/talks/blob/master/SQL_3/SQL_3.rmd

* dig up the RODBC reader https://github.com/mdsumner/mdsutils/blob/master/R/odbcReadManifold.R

* get rClr working with Manifold

* parallel CGAL triangulation code with Manifold's DecomposeToTrianglesAdv

### Manifold geometry via RODBC

We can read from Manifold map files using a bit of SQL and the wkb R package. 

```{r,eval=FALSE}
## extensions we need
library(wkb)    ## for parsing WKB blobs as Spatial R objects
library(sp)     ## Spatial R objects
library(RODBC)  ## ODBC in R
library(raster) ## just for nice print methods for sp objects
## pull in the odbcConnectManifold function
source("https://raw.githubusercontent.com/mdsumner/mdsutils/master/R/odbcReadManifold.R")

## open a connection to a map file
## remember, this file has Local Scale 0.0001
con <- odbcConnectManifold("E:\\DATA\\Manifold\\ManifoldCD\\Data\\World\\Medium Resolution\\World Provinces.map")
## list the available tables if needed
##sqlTables(con)

## read in just the ID and the Geom (I) as WKB 
## (Manifold's Geom includes the CRS so we cast to OGC using CGeomWKB)
## remember this is just a data.frame
ProvincesGeom <- sqlQuery(con, "SELECT [ID], [Country], [Province], CGeomWKB(Geom(ID)) AS [geom] FROM [Provinces] WHERE [Longitude (I)] > 100 AND [Latitude (I)] < 0")
## get the CRS (somehow)
## . . .
close(con)

## construct an R spatial object from the raw geometry
## this is just SpatialPolygons/Lines/Points (what happens to mixed geom layers?)
Rsp <- readWKB(ProvincesGeom$geom)

## reconstruct our original layer
## remember, this file has Local Scale 0.0001
Countries <- SpatialPolygonsDataFrame(Rsp, subset(ProvincesGeom, select = c("ID", "Country", "Province")))
Countries
#class       : SpatialPolygonsDataFrame 
#features    : 3453 
#extent      : 991389, 1800000, -783533, 20833  (xmin, xmax, ymin, ymax)
#coord. ref. : NA 
#variables   : 3
#names       :     ID, Country,                Province 
#min values  : 262126,        ,              Antarctica 
#max values  : 316877,  Tuvalu, Yogyakarta [Jogjakarta] 
devtools::session_info()
```

