---
title: "manifoldr: R for Manifold GIS"
author: "Michael D. Sumner"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Manifold GIS from R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# manifoldr



## Basic Usage


Read in a drawing. 

```{r}
library(manifoldr)
library(RODBC)
mapfile <- system.file("extdata", "AreaDrawing.map", package = "manifoldr")

geom2D <- DrawingA(mapfile, "Drawing")
geom2D

geom1D <- DrawingL(mapfile, "Drawing")
geom1D

geom0D <- DrawingP(mapfile, "Drawing")
geom0D

```

Unfortunately this must be done separately for points lines or areas, for now. Future versions will levarage `gris` or something like it to read a Drawing in its full state. 

Read in a surface. 


=======
```{r}
library(raster)
mapfile2 <- system.file("extdata", "Montara_20m.map", package= "manifoldr")

gg <- Surface(mapfile2, "Montara")

gg
```

Read in an image. 

```{r}
mapfile3 <- system.file("extdata", "V20160202016022.L3m_R3QL_NPP_CHL_chlor_a_9km.map", package= "manifoldr")
im <- Image(mapfile3, "V20160202016022.L3m_R3QL_NPP_CHL_chlor_a_9km")
plotRGB(im)
im
```

## Build a custom object with SQL


```{r}
mapfile <- system.file("extdata", "AreaDrawing.map", package = "manifoldr")

x <- manifoldr:::readmfd(mapfile, "Drawing",  query = "SELECT [Name], [Branches (I)], [X (I)] FROM [Drawing] WHERE isArea([ID]) ORDER BY [NAME]", spatial = FALSE)


```
## Lower level usage 

Open a connection to a built-in .map file and issue a query.  

```{r,eval=FALSE}
library(manifoldr)
library(RODBC)
mapfile <- system.file("extdata", "AreaDrawing.map", package = "manifoldr")
con <- odbcConnectManifold(mapfile)
tab <- sqlQuery(con, "SELECT [ID], [Name], BranchCount([ID]) AS [nBranch] FROM [Drawing] ORDER BY [nBranch]")
close(con)

print(tab)
```
=======
### Manifold geometry via RODBC

NOTE: there's no real need for this now, please use higher level functions which do all of this in one step. 


We can read from Manifold map files using a bit of SQL and the wkb R package. 

```{r,message=FALSE,warning=FALSE, eval=FALSE}
## extensions we need
library(wkb)    ## for parsing WKB blobs as Spatial R objects
library(sp)     ## Spatial R objects
library(RODBC)  ## ODBC in R
library(raster) ## just for nice print methods for sp objects
library(manifoldr)
## open a connection to a map file
## original  file has Local Scale 0.0001, so I use a modified copy "Provinces_"
con <- odbcConnectManifold("E:\\ManifoldDVD\\Data\\World\\Medium Resolution\\World Provinces.map")
## list the available tables if needed
##sqlTables(con)

## read in just the ID and the Geom (I) as WKB 
## (Manifold's Geom includes the CRS so we cast to OGC using CGeomWKB)
## remember this is just a data.frame
ProvincesGeom <- sqlQuery(con, "SELECT [ID], [Country], [Province], CGeomWKB(Geom(ID)) AS [geom] FROM [Provinces_] WHERE [Longitude (I)] > 100 AND [Latitude (I)] < 0")
## get the CRS (somehow)
## . . .
close(con)

## construct an R spatial object from the raw geometry
## this is just SpatialPolygons/Lines/Points (what happens to mixed geom layers?)
Rsp <- readWKB(ProvincesGeom$geom)

## reconstruct our original layer
Countries <- SpatialPolygonsDataFrame(Rsp, subset(ProvincesGeom, select = c("ID", "Country", "Province")))
Countries

plot(Countries)
devtools::session_info()
```
